Predictions From Day 0 Community Structures

### 2016-08-08 code/HumanCdGF_microbiota_predict.Rmd
Move echo=F to global opt

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
if(!require(randomForest)) install.packages('randomForest', repos = 'http://cran.us.r-project.org')
library(randomForest)
if(!require(miscTools)) install.packages('miscTools', repos = 'http://cran.us.r-project.org')
library(miscTools)
if(!require(ggplot2)) install.packages('ggplot2', repos = 'http://cran.us.r-project.org')
library(ggplot2)
```

```{r, echo=FALSE}
setwd("~/Documents/Github/Schubert_humanCdGF_XXXX_2016")

@@

#### main subset of meta file with all mouse and day 0 info
day_0_meta <- meta_file[which(meta_file$day==0),!is.na(meta_file['430-307-D0',])]
day_0_meta$RowNames <- rownames(day_0_meta)
day_0_meta$Euth_Early <-ifelse(day_0_meta$Euthanized %in% 10, 'No','Yes')
####
#    function to move mouse data from rows by mouse and days --> rows by mouse and columns by days
####
# select data to move
cdiff_effect <- c("cdiff_cfu","percent_weightLoss.x","Log_repiricoal_dilution")
# function input meta_file as DATAFRAME and mouse dataframe as INPUT
combine_days_by_mouse <- function(DATAFRAME, INPUT){
          OUTPUT <- INPUT
     for(DAY in 1:10){
          NEW_ROW <- DATAFRAME[which(DATAFRAME$day==DAY),c(cdiff_effect, 'Mouse_ID')]
          names(NEW_ROW)[!names(NEW_ROW) %in% 'Mouse_ID'] <- 
               paste('Day_',paste(DAY,paste('_',names(NEW_ROW)[!names(NEW_ROW) %in% 'Mouse_ID'], sep=''), sep=''),sep='')
          OUTPUT <- merge(OUTPUT, NEW_ROW, by='Mouse_ID', all.x=TRUE)
     }
     return(OUTPUT)
}

test <- merge(day_0_meta,meta_file[which(meta_file$day==3),c(cdiff_effect, 'Mouse_ID')], by='Mouse_ID')

rf_meta_df <- combine_days_by_mouse(meta_file[!meta_file$sample_type=='cecum',], day_0_meta)
# remove all columns with no data - toxin missing from some days
rf_meta_df <- rf_meta_df[,colSums(is.na(rf_meta_df)) != nrow(rf_meta_df)]
# add columns for data from day euthanized
day_euthanized_cdiff <- meta_file[which(meta_file$sample_type=='stool' & meta_file$day == meta_file$Euthanized), c(cdiff_effect,'Mouse_ID')]
names(day_euthanized_cdiff)[!names(day_euthanized_cdiff) %in% 'Mouse_ID'] <- 
     paste('Day_Euth_',names(day_euthanized_cdiff)[!names(day_euthanized_cdiff) %in% 'Mouse_ID'], sep='')
rf_meta_df <- merge(rf_meta_df,day_euthanized_cdiff, by='Mouse_ID', all.x=TRUE)

@@

```{r, echo=FALSE}
Day_Euth_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_Euth_cdiff_cfu')
Day_Euth_cdiff_cfu_rf$plot
```

@@

##Can we predict moribundity with the community structure at Day 0?
```{r, echo=FALSE}
##Fix to be categorical, and not regression
Predict_early_euth_df <- Predict_df[!Predict_df$Euth_Early %in% NA,c('Euth_Early',OTU_list)]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df)
rf_early_euth
```

@@

```{r, echo=FALSE}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot
```

@@

```{r,echo=FALSE}
#rf to predict 
Day_1_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_1_cdiff_cfu')
Day_1_cdiff_cfu_rf$plot
```

@@

# Day 2 C. difficile CFU
```{r, echo=FALSE}
Day_2_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_2_cdiff_cfu')
Day_2_cdiff_cfu_rf$plot
```

@@

# Day 3 C. difficile CFU
```{r, echo=FALSE}
Day_3_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_3_cdiff_cfu')
Day_3_cdiff_cfu_rf$plot
```

@@

# Day 4 C. difficile CFU
```{r, echo=FALSE}
Day_4_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_4_cdiff_cfu')
Day_4_cdiff_cfu_rf$plot
```

@@

# Day 5 C. difficile CFU
```{r, echo=FALSE}
Day_5_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_5_cdiff_cfu')
Day_5_cdiff_cfu_rf$plot
```

@@

# Day 6 C. difficile CFU
```{r, echo=FALSE}
Day_6_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_6_cdiff_cfu')
Day_6_cdiff_cfu_rf$plot
```

@@

# Day 7 C. difficile CFU
```{r, echo=FALSE}
Day_7_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_7_cdiff_cfu')
Day_7_cdiff_cfu_rf$plot
```

@@

# Day 8 C. difficile CFU
```{r, echo=FALSE}
Day_8_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_8_cdiff_cfu')
Day_8_cdiff_cfu_rf$plot
```

@@

# Day 9 C. difficile CFU
```{r, echo=FALSE}
Day_9_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_9_cdiff_cfu')
Day_9_cdiff_cfu_rf$plot
```

@@

# Day 10 C. difficile CFU
```{r, echo=FALSE}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot
```

@@

##seed greatly affects variance, likely due to data split, sample size might be too small to make reliable predictions

```{r knitr::kable, echo=FALSE, warning=F, message=F}
var_by_day <- data.frame(Day=c(1:10),Percent_Variance_Explained=NA)
cdiff_cfu <- vector()
for (a in 1:10){
          var_by_day$Percent_Variance_Explained[a] <- eval(parse(text = paste('Day_', a, '_cdiff_cfu_rf$rf$rsq[length(Day_', a, '_cdiff_cfu_rf$rf$rsq)]*100', sep='')))
          cdiff_cfu <- c(cdiff_cfu,paste('Day_', a, '_cdiff_cfu', sep=''))
     }

@@

##Can we predict day 1 weight change with the community structure at Day 0?
```{r, echo=FALSE}

@@

##Can we predict day 10 weight change with the community structure at Day 0?
```{r, echo=FALSE}

@@

##Can we predict ___________ with the community structure at Day 0?
```{r, echo=FALSE}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
if(!require(randomForest)) install.packages('randomForest', repos = 'http://cran.us.r-project.org')
library(randomForest)
if(!require(miscTools)) install.packages('miscTools', repos = 'http://cran.us.r-project.org')
library(miscTools)
if(!require(ggplot2)) install.packages('ggplot2', repos = 'http://cran.us.r-project.org')
library(ggplot2)
```

```{r}
setwd("~/Documents/Github/Schubert_humanCdGF_XXXX_2016")

@@

#### main subset of meta file with all mouse and day 0 info
day_0_meta <- meta_file[which(meta_file$day==0),!is.na(meta_file['430-307-D0',])]
day_0_meta$RowNames <- rownames(day_0_meta)
day_0_meta$Euth_Early <-ifelse(day_0_meta$Euthanized %in% 10, 'No','Yes')
####
#    function to move mouse data from rows by mouse and days --> rows by mouse and columns by days
####
# select data to move
cdiff_effect <- c("cdiff_cfu","percent_weightLoss.x","Log_repiricoal_dilution")
# function input meta_file as DATAFRAME and mouse dataframe as INPUT
combine_days_by_mouse <- function(DATAFRAME, INPUT){
          OUTPUT <- INPUT
     for(DAY in 1:10){
          NEW_ROW <- DATAFRAME[which(DATAFRAME$day==DAY),c(cdiff_effect, 'Mouse_ID')]
          names(NEW_ROW)[!names(NEW_ROW) %in% 'Mouse_ID'] <- 
               paste('Day_',paste(DAY,paste('_',names(NEW_ROW)[!names(NEW_ROW) %in% 'Mouse_ID'], sep=''), sep=''),sep='')
          OUTPUT <- merge(OUTPUT, NEW_ROW, by='Mouse_ID', all.x=TRUE)
     }
     return(OUTPUT)
}
rf_meta_df <- combine_days_by_mouse(meta_file[!meta_file$sample_type=='cecum',], day_0_meta)

# remove all columns with no data - toxin missing from some days
rf_meta_df <- rf_meta_df[,colSums(is.na(rf_meta_df)) != nrow(rf_meta_df)]
# add columns for data from day euthanized
day_euthanized_cdiff <- meta_file[which(meta_file$sample_type=='stool' & meta_file$day == meta_file$Euthanized), c(cdiff_effect,'Mouse_ID')]
names(day_euthanized_cdiff)[!names(day_euthanized_cdiff) %in% 'Mouse_ID'] <- 
     paste('Day_Euth_',names(day_euthanized_cdiff)[!names(day_euthanized_cdiff) %in% 'Mouse_ID'], sep='')
rf_meta_df <- merge(rf_meta_df,day_euthanized_cdiff, by='Mouse_ID', all.x=TRUE)

@@

```{r}
Day_Euth_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_Euth_cdiff_cfu')
Day_Euth_cdiff_cfu_rf$plot
```

@@

##Can we predict moribundity with the community structure at Day 0?
```{r}
##Fix to be categorical, and not regression
Predict_early_euth_df <- Predict_df[!Predict_df$Euth_Early %in% NA,c('Euth_Early',OTU_list)]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df)
rf_early_euth
```

@@

```{r}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot
```

@@

```{r}
#rf to predict 
Day_1_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_1_cdiff_cfu')
Day_1_cdiff_cfu_rf$plot
```

@@

# Day 2 C. difficile CFU
```{r}
Day_2_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_2_cdiff_cfu')
Day_2_cdiff_cfu_rf$plot
```

@@

# Day 3 C. difficile CFU
```{r}
Day_3_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_3_cdiff_cfu')
Day_3_cdiff_cfu_rf$plot
```

@@

# Day 4 C. difficile CFU
```{r}
Day_4_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_4_cdiff_cfu')
Day_4_cdiff_cfu_rf$plot
```

@@

# Day 5 C. difficile CFU
```{r}
Day_5_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_5_cdiff_cfu')
Day_5_cdiff_cfu_rf$plot
```

@@

# Day 6 C. difficile CFU
```{r}
Day_6_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_6_cdiff_cfu')
Day_6_cdiff_cfu_rf$plot
```

@@

# Day 7 C. difficile CFU
```{r}
Day_7_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_7_cdiff_cfu')
Day_7_cdiff_cfu_rf$plot
```

@@

# Day 8 C. difficile CFU
```{r}
Day_8_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_8_cdiff_cfu')
Day_8_cdiff_cfu_rf$plot
```

@@

# Day 9 C. difficile CFU
```{r}
Day_9_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_9_cdiff_cfu')
Day_9_cdiff_cfu_rf$plot
```

@@

# Day 10 C. difficile CFU
```{r}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot
```

@@

## seed greatly affects variance, likely due to data split, sample size might be too small to make reliable predictions

```{r knitr::kable, warning=F, message=F}
var_by_day <- data.frame(Day=c(1:10),Percent_Variance_Explained=NA)
cdiff_cfu <- vector()
for (a in 1:10){
          var_by_day$Percent_Variance_Explained[a] <- eval(parse(text = paste('Day_', a, '_cdiff_cfu_rf$rf$rsq[length(Day_', a, '_cdiff_cfu_rf$rf$rsq)]*100', sep='')))
          cdiff_cfu <- c(cdiff_cfu,paste('Day_', a, '_cdiff_cfu', sep=''))
     }

@@

##Can we predict day 1 weight change with the community structure at Day 0?
```{r}

@@

##Can we predict day 10 weight change with the community structure at Day 0?
```{r}

@@

##Can we predict ___________ with the community structure at Day 0?
```{r}
